### 2021-09-03 input/figadd.tex
cosmetic + number
first step in styling
ongoing wip

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:addA}Both \Process~$a$ and \Process~$d$
      become sources.  $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]  {\input{input/figaddA.tex}}
    \hspace{5pt}
    \subfloat[Part B][\label{fig:addB}Messages transit through communication
      links and carry increasing weights.]{\input{input/figaddB.tex}}
    \hspace{5pt} \subfloat[Part C][\label{fig:addC}\Process~$b$ and
      \Process~$c$ receive, deliver, and forward $\alpha_{d}^{1}$ and
      $\alpha_d^3$ respectively.]{\input{input/figaddC.tex}}
    \hspace{5pt} \subfloat[Part D][\label{fig:addD}$a$ and
      $b$ discarded their received messages. $c$ still improved with $\alpha_d^2$.]
    {\input{input/figaddD.tex}}
    \caption{\label{fig:add}Efficient consistent partitioning using
      \NAMEB. Partition~$P_a$ includes $a$ while Partition~$P_d$
      includes $b$, $c$, and $d$. \Process~$c$ and \Process~$d$ never
      acknowledge the existence of Source~$a$, for \Process~$b$ stops
      the propagation of the latter's notifications.}
  \end{center}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:addA}Both $a$ and $d$
      become sources.  $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]  {\input{input/figaddA.tex}}
    \hspace{1pt}
    \subfloat[Part B][\label{fig:addB}Messages transit through %communication
      links and carry increasing weights.]{\input{input/figaddB.tex}}
    \hspace{1pt}
    \subfloat[Part C][\label{fig:addC}$b$ and
      $c$ receive, deliver, and forward $\alpha_{d}^{1}$ and
      $\alpha_d^3$ respectively.]{\input{input/figaddC.tex}}
    \hspace{1pt}
    \subfloat[Part D][\label{fig:addD}$a$ and
      $b$ discarded their received messages. $c$ still improved with $\alpha_d^2$.]
    {\input{input/figaddD.tex}}
    \caption{\label{fig:add}Efficient consistent partitioning using
      \NAMEB. Partition~$P_a$ includes $a$ while Partition~$P_d$
      includes $b$, $c$, and $d$. \Process~$c$ and \Process~$d$ never
      acknowledge the existence of Source~$a$, for \Process~$b$ stops
      the propagation of the latter's notifications.}
  \end{center}
\end{figure*}

### 2021-09-03 input/figproblem.tex
first step in styling
ongoing wip

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both \Process~$a$ and
      \Process~$d$ add then delete their partition. For the recall,
      $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{1pt}
    \subfloat[Part B][\label{fig:problemB}\Process~$b$
      delivers and forwards the $\alpha$ message about $P_a$ while
      \Process~$c$ delivers and forwards both $\alpha$ and $\delta$
      messages about $P_d$.]  {\input{input/figproblemB.tex}}
    \hspace{1pt}
    \subfloat[Part C][\label{fig:problemC}\Process~$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before. \Process~$c$ does not deliver
      $\alpha_d$, for it already deleted it
      before.]{\input{input/figproblemC.tex}}
    \hspace{1pt}
    \subfloat[Part D][\label{fig:problemD}\Processes
      discard remaining messages. Since \Process~$b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted partition
      $P_a$.]{\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:problemA}Both \Process~$a$ and
      \Process~$d$ add then delete their partition. For the recall,
      $w_{ab} = 2$; $w_{bc} = w_{bd} = 1$; $w_{cd} =
      3$.]{\input{input/figproblemA.tex}}
    \hspace{0pt}
    \subfloat[Part B][\label{fig:problemB}$b$
      delivers and forwards the $\alpha$ message about $P_a$ while
      \Process~$c$ delivers and forwards both $\alpha$ and $\delta$
      messages about $P_d$.]  {\input{input/figproblemB.tex}}
    \hspace{0pt}
    \subfloat[Part C][\label{fig:problemC}$b$
      already blocks the propagation of $\delta_a$ despite it
      propagated $\alpha_a$ before; and $c$ does not deliver
      $\alpha_d$, for it already deleted it
      before.]{\input{input/figproblemC.tex}}
    \hspace{0pt}
    \subfloat[Part D][\label{fig:problemD}\Processes
      discard remaining messages. Since $b$ cannot deliver
      $\delta_a$, $c$ and $d$ remain in the deleted $P_a$.]
             {\input{input/figproblemD.tex}}
  \end{center}
  \caption{\label{fig:problem}Naive propagation of $\alpha$ and
    $\delta$ messages is insufficient to guarantee consistent partitioning.}
\end{figure*}

### 2021-09-03 input/proposal.tex
ongoing wip

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: {\small{$b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$}}.
\end{definition}

@@

\begin{definition}[\label{def:lwo}Last win order and staleness]
  When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  $\nexists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m$
  convey \emph{stale} control information.
\end{definition}

@@

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending. Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ does not deliver $\alpha_d$ and $\delta_d$ since it
already delivered it. The system converges to an inconsistent state
where some \processes assume they belong to a partition while it does
not exist anymore. Naive propagation of $\alpha$ and $\delta$ messages
is insufficient to guarantee consistent partitioning when any \process
can add or delete its partition at any time. \Processes need to
exploit their local knowledge gathered over message receipts.

@@

\begin{theorem}[\label{theo:dcp}purge + BEF $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When each \process can \texttt{Add} and \texttt{Del}ete its partition
at any time, consistent partitioning requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($b_p(m) \rightarrow b_p(m')
  \implies \eventually (d_q(m) \implies d_q(m) \rightarrow
  d_q(m''))$);
\item and best eventual forwarding.
\end{inparaenum}
\end{theorem}

\begin{proof}
  \Processes deliver and forward the best sum of weights they
  received. Figure~\ref{fig:proofA} depicts the only case where a
  \process would block the epidemic propagation: the \process received
  and delivered a message from a partition that has since then been
  deleted but with a better sum of weights than newly received ones
  (in the example $\alpha_d^y < \alpha_a^{x'}$). Removing
  \emph{forever} all such stale messages about deleted partitions
  would allow \processes to propagate their best up-to-date partition
  again, eventually reaching consistent partitioning together, as
  stated by Theorem~\ref{theo:bef}.
\end{proof}

@@

\begin{theorem}[three-phase purge] % Propagations $+$ detection $\implies$ purge]
  Purging stale control information from its source to every \process
  that may have delivered it last requires three
  phases:
  \begin{inparaenum}[(i)]
  \item propagation of delete notifications,
  \item detection of possible blocking conditions, and
  \item propagation of possibly deleted but blocked notifications.
  \end{inparaenum}
\end{theorem}

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source. Each message $m$ carries its source $s$ along with its version
$c$ that we subscript $m_{s, c}$.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  Its size
increases linearly with the number of sources that the \process has
ever known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

>>>>

\begin{definition}[\label{def:scoped}\underline{S}coped broad\underline{cast} (\NAMEB)]
  When \Process~$x$ scoped \underline{b}roadcasts $b_x(m)$ a
  \underline{m}essage $m$, every correct \process $y$ within a scope
  \underline{r}eceives $r_y(m)$ and \underline{d}elivers it
  $d_y(m)$. The scope depends on the \underline{s}tate $\sigma$ of
  each \process, the \underline{m}etadata $\mu$ piggybacked by each
  message, and a \underline{p}redicate $\phi$ verified from \process to
  \process: $b_x(m) \wedge r_y(m) \implies \exists \pi_{xy}: \forall z
  \in \pi_{xy}, \phi(\mu_z, \sigma_z)$.
\end{definition}

@@

\begin{definition}[\label{def:lwo}Message staleness]
  A message $m$ conveys \underline{s}tale control information if the
  process that broadcast $m$ later broadcast another message $m'$:
  $S(m) = \exists m': b_p(m) \rightarrow b_p(m')$.

  %% When a \process $p$ broadcasts two messages $b_p(m) \rightarrow
  %% b_p(m')$, no \process $q$ can deliver $m$ if it has delivered $m'$:
  %% $\nexists q \in V$ with $d_q(m') \rightarrow d_q(m)$, for $m$
  %% convey \emph{\underline{s}tale} control information: $S(m)$.
\end{definition}

@@

Unfortunately, stale control information as stated in
Definition~\ref{def:lwo} may impair the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} highlights such consistency issues with
dynamic partitions, even in contexts where \processes have FIFO links,
\ie where \processes receive the messages in the order of their
sending ($s_{pq}(m) \rightarrow s_{pq}(m') \implies r_q(m) \rightarrow
r_q(m')$). Both \Process~$a$ and \Process~$d$ add then delete their
partition concurrently. \Process~$c$ receives, delivers, and forwards
$\alpha_d^3$ followed by $\delta_d$. In the meantime, \Process~$b$
receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, both \Process~$c$ and \Process~$d$ deliver
the message about Partition $P_a$, for they did not have any known
partition at receipt time. On the contrary, \Process~$b$ delivers
$\alpha_d^1$, for it improves its distance to a known
source. \Process~$b$ then blocks \Process~$a$'s removal notification
$\delta_a$. It never reaches \Process~$c$ nor \Process~$d$. Also,
\Process~$c$ does not deliver $\alpha_d$ and $\delta_d$ since it
already delivered it. The system converges to an inconsistent state
where some \processes assume they belong to a partition while it does
not exist anymore. Naive propagation of $\alpha$ and $\delta$ messages
is insufficient to guarantee consistent partitioning when any \process
can add or delete its partition at any time. \Processes need to
exploit their local knowledge gathered over message receipts.

@@

\begin{theorem}[\label{theo:dcp}Purge + BEF $\implies$
    \underline{D}ynamic \underline{c}onsistent
    \underline{p}artitioning (DCP)]
%
When each \process can \texttt{Add} and \texttt{Del}ete its partition
at any time, consistent partitioning requires
\begin{inparaenum}[(i)]
\item eventual purging of stale messages ($S(m)
  \implies (d_q(m) \implies \eventually d_q(m) \rightarrow
  d_q(m'))$);
\item and best eventual forwarding so such an $m'$ ends up originating from 
  the best up-to-date source for $q$.
\end{inparaenum}
\end{theorem}

\begin{proof}
  %% To reach consistent partitioning, the last message delivered by
  %% every process is not stale and comes from its respective best
  %% source: $d(m)$
  % As for CP, DCP requires that every process' last delivery

  \processes deliver and forward the best sum of weights they
  received. Figure~\ref{fig:proofA} depicts the generic case where a
  \process would block the epidemic propagation: the \process received
  and delivered a message from a partition that has since then been
  deleted but with a better sum of weights than newly received ones
  (in the example $\alpha_d^y < \alpha_a^{x'}$). Removing
  \emph{forever} all such stale messages about deleted partitions
  would allow \processes to propagate their best up-to-date partition
  again, eventually reaching consistent partitioning together, as
  stated by Theorem~\ref{theo:bef}.
\end{proof}

@@

\begin{theorem}[Three-phase purge] % Propagations $+$ detection $\implies$ purge]
  Purging stale control information from its source to every \process
  that may have delivered it last requires three
  phases:
  \begin{inparaenum}[(i)]
  \item propagation of delete notifications,
  \item detection of possible blocking conditions, and
  \item propagation of possibly deleted but blocked notifications.
  \end{inparaenum}
\end{theorem}

@@

To implement the removal of stale messages as stated in
Definition~\ref{def:lwo}, each \process maintains a vector of versions
that associates the respective known local counter of each known
source, or has-been source. Each message $m$ carries its source $s$
along with its version $c$ that we subscript $m_{s, c}$.  Each
\process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  Its size
increases linearly with the number of sources that the \process has
ever known, which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.
